---
title: 'Workshop 3.1: Visualise with ggplot2'
author: "James Morgan, Lucy Liu"
date: '`r format(Sys.time(), "Last modified: %d %b %Y")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install the packages

Option 1: Install the entire tidyverse  
This option will install ggplot, dplyr and about 70 other packages that are designed to work together under the same philosophy of tidy data.

```{r Install the tidyverse, eval=FALSE}
install.packages("tidyverse")
```

Option 2: Install ggplot2 only
```{r Install ggplot2, eval=FALSE}
install.packages("ggplot2")
```

## Load the packages

Packages are an extension on the capabilities of Base R. A package may be installed, but it can only be used after explicitly loading it into our envirionment.  
```{r Load ggplot2, eval=FALSE}
library(ggplot2)
```

## Visualise!
For this session, we'll be using a built in dataset: "Miles per gallon"

Before we 
```{r Check that data}
mpg
```

```{r Your turn #1}
ggplot(data=mpg) +
    geom_point(mapping=aes(x=displ,y=hwy)) +
    geom_smooth(mapping=aes(x=displ,y=hwy), method = 'lm')
```

```{r}

```

