---
title: "Workshop 3.1: Visualise with ggplot2"
author: James Morgan, Lucy Liu, Hadley Wickham
date: '`r format(Sys.time(), "Last modified: %d %b %Y")`'
output: html_document
---

## Package Loading

Packages are an extension on the capabilities of Base R. A package may be installed, but it can only be used after explicitly loading it into our envirionment.  
```{r setup}
library(ggplot2)
```

## Check your Data
For this session, we'll be using a dataset that's provided with ggplot2: "mpg"

Before we start to visualise our data, you should always know a little bit about it. So let's take a look!.


```{r Check that data}
#View(mpg)
mpg
```

Check the data structure.
```{r Check that data 2}
str(mpg)
```

Generate a summary of the data inside the data frame.
```{r Check that data 3}
summary(mpg)
```

Check the documentation...
```{r Check that data 4, eval=FALSE}
?mpg
```
Note that this one only works because this dataset is built in and documented by the package maintainers.

## Quiz

What relationship do you expect to see between engine size (displ) and highway fuel efficiency (hwy)?

----------------------------------------------------------------------

## ggplot and The Grammar of Graphics

### ggspele
ggplot is one of the most popular R packages. It allows you to create good looking graphs quickly and easily. It builds on the concept of the “Grammar of Graphics” (Wilkinson 2005, Bertin 1983) which describes a consistent syntax for the construction of a wide range of complex graphics by a concise description of their components.

The structured syntax and high level of abstraction used by ggplot2 should allow for the user to concentrate on the visualisations instead of creating the underlying code.

On top of this central philosophy ggplot2 has:

* Increased flexibility over many plotting systems.
* An advanced theme system for professional/publication level graphics.
* Large developer base – Many libraries extending its flexibility.
* Large user base – Great documentation and active mailing list.

Let's start with the basic anatomy of ggplot code - or how to give R instructions to make your plot. 
First you must tell ggplot what data you want to plot. 

Ultimately - The "grammar of graphics" works by introducing consecutive layers of content onto a plot.

First, we initalise the plot. `ggplot()` creates a plot object and links it to a dataset.
```{r Intro to ggplot}
ggplot(data = mpg)
```

## Your Turn 1

Type in the code from the slide, and run it to make a graph. Pay strict attention to spelling, capitalization, and parentheses.

I strongly recommend typing it, even though copying and pasting is easier! Re-typing it will help your fingers learn the patterns you'll need to apply to your own data analyses.

```{r}


```

Remember to raise a post-it if you get stuck, and then once you're done!

## Your Turn 2

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

## Your Turn 3

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Guess!

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy))
```

## Your Turn 4

Make a histogram of the `hwy` variable from `mpg`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 5

Make a density plot of `hwy` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 6

Make a bar chart `class` colored by `class`.

```{r}
ggplot(data = mpg) 

```

## Your Turn 7

Predict what this code will do. Then run it.

```{r}
ggplot(mpg) + 
  geom_point(aes(displ, hwy)) +
  geom_smooth(aes(displ, hwy))
```

## Your Turn 8

What does `getwd()` return?

```{r}
getwd()
```

## Your Turn 9

Save the last plot and then locate it in the files pane. Next find it on your desktop.

```{r}

```

***

# Take aways

You can use this code template to make thousands of graphs with **ggplot2**.

```{r, eval = FALSE}
ggplot(data = <DATA>) +
    <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>)) +
    <COORDINATE_FUNCTION> +
    <FACET_FUNCTION> +
    <SCALE_FUNCTION> +
    <THEME_FUNCTION>
```

