---
title: "Workshop #3 Exercises: dplyr & ggplot"
author: Lucy Liu, James Morgan
output: html_notebook
---

The contents of this workshop are adapted from the University of Cambirdge [Bioinformatics training](https://bioinfotraining.bio.cam.ac.uk/) workshop ["Data Manipulation and Visualisation using R"](http://bioinformatics-core-shared-training.github.io/r-intermediate/)

```{r setup}
#setwd('your/filepath/here/r-intro/Day-3')

patients <- read.csv("patients-info.txt", sep="\t")

library(tidyverse)
```

#Data Exploration
```{r}
head(patients,10)

str(patients)
    #have a think about what format you'd expect these variables to be. What do you think about `Birth`? 
```

###Return a data frame that contains only those individuals who have a pet Dog or Bird
```{r}

```

### Using the same criteria, show only `Name`, `Race` & `Sex`.
```{r}

```

### Save a data frame with an extra column that contains the BMI for each individual.

```{r}

```

### Return a data frame containing the name, sex, age and weight for all overweight (BMI>25) individuals who smoke, and are still alive. 

```{r}

```

### Return a data frame containing the average patient height.
```{r}

```
### Now report the average weight per state.
```{r}

```


### Can you report average heights and weights for males and females in each state??
```{r}

```

### Challenge: Find the proportion of smokers in each sex?
    Hint - R stores and can interpret booleans as follows: TRUE == 1, FALSE == 0. 
    So we can calculate a proportion by taking the mean of a boolean vector.
```{r}

```
